---
title: "Preliminary Analysis"
author: "Eduardo Ariño de la Rubia"
date: "December 25, 2014"
output: html_document
---

So this is my first EDA on this new data set, I downloaded the data from
killedbypolice.net and after some wrangling, I think I've turned it into 
an actual usable data set. For what yet, I dont' know.

At the end of this knitr document i create some csv files and a graph that
shows that the north east is way less eager to kill population than the south
west.

data sets:

* [full cleaned data](data/cleaned.csv)
* [summary by state](data/summaries_by_state.csv)
* [summary_by_state_and_race](data/summaries_by_state_and_race.csv)

```{r}
library(tidyr)
library(dplyr)
library(lubridate)
library(Hmisc)
library(ggplot2)
library(maps)
library(rvest)
library(httr)

data(state)
states_map <- map_data("state")
```


```{r}
# index.tidy.html was created using http://infohound.net/tidy/ from
# killedbypolice.net after some small manual editing (mainly deleting
# stuff that I knew would cause parsing headaches)
#
# tidy data/index.html > data/index.tidy.html

columns <- c("date", "state", "gender_race", "name_age", "kbp_link", "news_link")
killed_by_police <- html("data/index.tidy.html")
data <- killed_by_police %>% html_table(header=TRUE) %>% 
  .[[1]] %>%
  setNames(columns) %>%
  mutate(#first thing we have to do is split up columns w/multiple values
    gender=unlist(lapply(strsplit(as.character(gender_race), "\\/"), "[", 1)), 
    race=unlist(lapply(strsplit(as.character(gender_race), "\\/"), "[", 2)),
    name=unlist(lapply(strsplit(as.character(name_age), "\\,"), "[", 1)),
    age=unlist(lapply(strsplit(as.character(name_age), "\\, "), "[", 2)),
    first_name=unlist(lapply(strsplit(as.character(name), " "), "[", 1)),
    #throw out some cruft from some strings
    gender=substr(gender, 1, 1),
    race=substr(race, 1, 1),
    date=gsub("\\([^)]+) ", "", date),
    #make a real date object
    date=mdy(date),
    age=as.numeric(substr(age, 1, 2)),#centegenerians will be turned into teens
    gender=as.factor(gender),
    race=as.factor(race),
    first_name=as.factor(capitalize(tolower(first_name))),
    decade=as.factor(age %% 10),
    state=as.factor(state.name[match(state, state.abb)]))

write.csv(data, "data/cleaned.csv", row.names=FALSE)

state_population <- read.csv("data/state_population.csv") %>%
  filter(NAME %in% state.name)

summaries_by_state <- data %>% group_by(state) %>% 
  summarise(total=n(), 
            mean_age=round(mean(age, na.rm=TRUE))) %>% 
  inner_join(state_population, by=c("state" = "NAME")) %>% 
  select(state, mean_age, total, POPESTIMATE2014, 
         POPEST18PLUS2014) %>% 
  mutate(pctg=(total/POPESTIMATE2014)*100) %>% 
  arrange(state) %>%
  mutate(state=tolower(state)) %>% mutate(state_rank=rank(pctg))

write.csv(summaries_by_state, "data/summaries_by_state.csv", row.names=FALSE)

summaries_by_state_and_race <- data %>% group_by(state, race) %>% 
  summarise(total=n(), 
            mean_age=round(mean(age, na.rm=TRUE))) %>% 
  inner_join(state_population, by=c("state" = "NAME")) %>% 
  select(state, race, mean_age, total, POPESTIMATE2014, 
         POPEST18PLUS2014) %>% 
  mutate(pctg=(total/POPESTIMATE2014)*100) %>% 
  arrange(state) %>%
  mutate(region=state.region[match(state.name[state], state.name)])

write.csv(summaries_by_state_and_race, "data/summaries_by_state_and_race.csv", row.names=FALSE)

pctg_map <- summaries_by_state %>% 
  inner_join(states_map, by=c("state" = "region")) %>% 
  arrange(group, order)

ggplot(pctg_map, aes(x=long, y=lat, group=group, fill=state_rank)) +
  geom_polygon(colour="black") + 
  scale_fill_gradient2(low="#559999", mid="grey90", high="#BB650B", 
                       midpoint=median(pctg_map$state_rank)) + 
  coord_map("polyconic") + ggtitle("Ranking Percentage of Population Killed by Police by State")

data <- data %>% filter(date > ymd("2014/01/01"))
```

You can also embed plots, for example:

```{r, echo=FALSE}
ggplot(data, aes(x=date)) + geom_density() + facet_wrap(~ race)

j <- data %>% group_by(first_name, race) %>% summarise(killed=n()) %>% arrange(killed)

data.frame(j) %>% arrange(killed) -> names_by_killed
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
